---
title: "plot_geolocated_points_in_map"
author: "Rubén Rellán Álvarez"
date: "July 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
```{r}
library(tidyr)
library(plyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(rmarkdown)
library(cowplot)
library(ggmap)
library(maps)
```

## Plotting geolocated points into a map using `ggmap`


```{r}
#load sample data

andosol <- read.csv("data/andosol.csv")

#create variables from data 

longitude <- andosol$Long
latitude <- andosol$Lat
id <- andosol$ID

#create basemap using get_map function
#see more detailed info about get_map here: http://www.r-bloggers.com/gps-basemaps-in-r-using-get_map/
# play with different levels of zoom and maptypes.

basemap <- get_map(location='Guanajuato, Mexico', zoom = 5, maptype = "satellite", color='bw', source='google')
ggmap(basemap)

#save the base layer map. Notice that I am adding a 

map <- ggmap(basemap, extent='panel', base_layer=ggplot(andosol, aes(x=longitude, y=latitude)))

#add points

map + geom_point(aes(colour = P_retention)) + 
  ggsave("map_races_p_retention.png", height = 12, width = 12, units = "cm")
```


